---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# btransfer

<!-- badges: start -->
<!-- badges: end -->

The goal of `btransfer` is to provide useful functions to perform unit value transfer
with income adjustment between countries or country aggregates. It also provides
a function to compute the social discount rate for any given country or country
aggregates.

## Installation

You can install the development version of `btransfer` from [GitHub](https://github.com/) with:

``` {r install, eval = F}
# install.packages("devtools")
devtools::install_github("giuliomela/btransfer")
```

## Examples

The main function of the package is `btransfer` which computes so-called transfer
factors to perform unit value transfer from a user-defined study site to a policy 
site. The user can choose the place and year in which the original study was carried out
(with `study_site` and `study_yr` options) as well as the place and year to which
the underlying value has to be transferred (`policy_site` and `policy_yr`).
The function also considers inflation and exchange rates when performing the
computation. The user can define both study and policy currencies (`study_currency` and
`policy_currency`), while the adjustment for inflation is always carried out using
the GDP deflator of the `policy_currency`. Transfer factors can be computed also when
the policy site is an aggregate. The user can control this option with the
argument `agg_policy`. The transfer factor can be computed also for a year in the future
up to 2050.

```{r example, echo=TRUE, eval=FALSE}

study_site <- "United Kingdom"

policy_site <- c("Italia", "France", "Deutschland", "Polska", "United States")

# Value transfer to the EU to some individual countries. No currency conversion.

btransfer::btransfer(study_site = "Unione Europea", policy_site = "Italy", 
            study_yr = 2016, policy_yr = 2019, ref_yr = 2019, policy_currency = "EMU",
            study_currency = "EMU")

# The primary value is expressed in USD, but must be converted to euro

btransfer::btransfer(study_site = "Australia", policy_site = "Italy", 
            study_yr = 2014, policy_yr = 2020, ref_yr = 2019, study_currency = "USA",
            policy_currency = "EMU")

# The primary value is expressed in the study country's currency and must be
# converted to the currency of the policy country

btransfer::btransfer(study_site = "Mexico", policy_site = "India", 
            study_yr = 2014, policy_yr = 2020, ref_yr = 2019, study_currency = "MEX",
            policy_currency = "IND")

# The study site is the world as a whole. Original estimate in USD, but value
# must be converted to euro.

# defining the countries that make up the aggregate
nos <- c("France", "United Kingdom", "Belgium", "Netherlands",
         "Denmark", "Norway", "Germany") # North Sea countries

btransfer::btransfer(study_site = "World", policy_site = nos, 
            study_yr = 2014, policy_yr = 2020, ref_yr = 2019, study_currency = "USA",
            policy_currency = "EMU", agg_policy = "yes")

# The maritime aggregates included in the EU's Handbook on the External Cost of Transport
# are pre-loaded in the package and can be directly selected as either study or policy
# site. See the help page of `btransfer` for more information

# The policy year is in the future

btransfer::btransfer(study_site = "World", policy_site = "Italy", 
            study_yr = 2014, policy_yr = 2030, ref_yr = 2019, study_currency = "USA",
            policy_currency = "EMU", agg_policy = "no")

```

Another useful function provided by the package is `compute_sdr` which calculates
the social discount rate for selected countries or country aggregates using the
Ramsey's rule. This parameter is useful in economic Cost Benefit Analysis or other
economic evaluations in which monetary values must be discounted back from the
future to present in order to perform the assessment.

```{r example-sdr, echo=TRUE, results='hide', eval=FALSE}

# The SDR calculated for a single country

btransfer::compute_sdr(country = "Italy", policy_yr = 2019, ref_yr = 2019, h = 10)

# SDR calculated for the World as a whole

btransfer::compute_sdr(country = "World", policy_yr = 2019, ref_yr = 2019, h = 10)

# SDR calculated for an aggregate of countries

btransfer::compute_sdr(country = nos, policy_yr = 2019, ref_yr = 2019, h = 10,
            aggregate = "yes", aggregate_name = "North Sea")


# SDR calculated for the rest of the world, given an aggregate of countries

btransfer::compute_sdr(country = nos, policy_yr = 2019, ref_yr = 2019, h = 10,
            aggregate = "row")

```

