---
title: "btransfer"
author: "Giulio Mela"
date: "`r format(Sys.time(), '%d - %m - %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{btransfer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: value_transfer.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=FALSE}
devtools::load_all(".")
```

# Value transfer `btransfer`

## Introduction {#intro}

The package `btransfer` allows to automatize unit value transfer with income
adjustment, a technique widely used in environmental and health economics to estimate
the value of an environmental or health good or service at a so-called _policy site_
by assigning to it the value of a similar good/service elsewhere, at a _study site_.
Value transfer methods can be of many different types. This package focuses on the
most widely used in environmental economics literature: unit value transfer with 
income adjustment.
This technique allows to adjust the original estimate taking into account differences
in income (and therefore also in willingness to pay for the good) between study and
policy sites. More sophisticated methods - such as function transfer - take into
consideration a greater number of parameters but are more complicated and
data-intensive and do not always yield better results.
The unit value transfer with income adjustment can be performed using the following
formula:

$$
V_{pi} = V_{sj}\left(\frac{V_{sj}}{Y_{sj}}\right)^{\epsilon}
$$

where $V_{pi}$ is the value at the policy site $p$ in year $i$, $V_{sj}$ is the
value to be transferred at the study site $s$ in year $j$, $Y_{sj}$ is the gross
domestic product (GDP) per capita in the study site in year $j$, $V_{pi}$ is the
GDP per capita in the policy site in year $i$ and $\epsilon$ is the income
elasticity of the willingness to pay for the good.
GDP per capita is expressed at purchasing power parities and in constant terms to
consider differences in the general price levels across countries [@Brander2013].
The demand for most goods and services is a positive function of income: as the 
latter rises, the former rises accordingly [@Brander2013]. However, such increase is not always
proportional since, at different income levels, individuals' preferences may change.
$\epsilon$ can therefore assume different values, depending on income levels.
@navrud2007a and @Navrud2009 provide insights on the relationship between $\epsilon$
and income levels: the elasticity of willingness to pay generally decreases as
income increase as for many normal goods. Following @Navrud2009, $\epsilon$ is assumed
to be $0.2$ and $0.5$ for high and upper-middle income countries, following the
World Bank income classification of countries. For lower-middle and low income
countries - following an email exchange with the same Stale Navrud - $\epsilon$ 
values of $0.8$ and $1.2$ are assumed.
Income differences between policy and study sites are not the only variables that
must be considered when performing a unit value transfer, since primary estimates
at the study site are usually reported in the study site currency and at the general
price level for a particular year [@Brander2013]. Therefore, after performing the
value transfer proper, values must be adjusted for inflation using the GDP deflator
and, if needed, also by applying appropriate nominal exchange rates.
The `btransfer` package can perform all these operations for almost any country
in the world as well as aggregates of countries also for years in the future
(up to 2050). It can be installed from GitHub as follows:

```{r install, eval = FALSE}

# Install from GitHub

devtools::install_github("https://github.com/giuliomela/btransfer")

```


The following sections explain the typical workflow.

## Identify study and policy sites {#study-policy}

The first step of the analysis is to identify from which country the primary
estimate comes from (study site) and the country(ies) to which such estimate
must be transferred.

```{r seelcting-countries}

study_site <- "United Kingdom"

policy_site <- c("Italia", "France", "Deutschland", "Polska", "United States")

```

Both `study_site` and `policy_site` can be expressed in any language as `btransfer`
automatically detects it. `study_site` must be a single string, while `policy_site`
can also be a vector.
In case the `study_site` is the world as a whole (i.e. like in the case of  
greenhouse gas emissions) or the European Union, special strings must be used:
`WLD` and `EUU` respectively.

## Usage of `bt_transfer`

The main function of the package is `bt_transfer`, which computes transfer factors
to be multiplied by value that must be transferred in order to perform the value
transfer. Such factors incorporate the value transfer proper as well as adjustments
for inflation and exchange rates.

```{r example-base, echo=TRUE}

# Value transfer to the EU to some individual countries. No currency conversion.

bt_transfer(study_site = "EUU", policy_site = policy_site, 
            study_yr = 2016, policy_yr = 2019, policy_currency = "EUR",
            study_currency = "EUR")

```

The function `bt_transfer` returns a tibble with the key variables used to calculate
the transfer factor and the transfer factor itself. Other than study and policy
sites, the user must provide also the year in which the study was carried out 
(`study_yr`) and the year to which the value must be transferred (`policy_yr`).
Default values for `study_yr` and `policy_yr` are 2016 and 2019 respectively.
By default, `bt_transfer` assumes that the value that must be transferred is
already expressed in euro. In case it is not, alternative currencies can be
supplied by the user using the arguments `study_currency` and `policy_currency`.
Other than the default `EUR`, the user can select US dollars (`USD`) or the currency
of study and policy sites (`LCU`). This latter option, however, can be used only
if just one policy site is selected.

```{r example-currency, echo=TRUE}

# The primary value is expressed in USD, but must be converted to euro

bt_transfer(study_site = "Australia", policy_site = policy_site, 
            study_yr = 2014, policy_yr = 2020, study_currency = "USD",
            policy_currency = "EUR")

```

```{r example-currency2, echo=TRUE}

# The primary value is expressed in the study country's currency and must be
# converted to the currency of the policy country

bt_transfer(study_site = "Mexico", policy_site = "India", 
            study_yr = 2014, policy_yr = 2020, study_currency = "LCU",
            policy_currency = "LCU")

```


Both the `study_site` and the `policy_site` parameters can be set to `WLD`, if
the study and/or the policy sites are the world as a whole. This option is
useful especially when one has to transfer cost factors related to externalities
at global level, such as the impact of air pollution on the climate.
Other than for individual countries, transfer factors can be calculated also for
aggregates of countries, using the parameter `aggregate`. Such parameter can assume
three values: `no` (the default), `yes` and `row`. In case `yes` is selected,
the transfer factor is calculated for the aggregate made up by the list of 
countries supplied through `policy_site`. Conversely, when `row` is selected,
the transfer factor is calculated for the rest-of-the-world aggregate obtained
subtracting the selected countries from the world total. When `aggregate` is set
to either `yes` or `row`, the parameter `policy_currency` can be set to `EUR` or
`USD` only.

```{r example-aggregates, echo=TRUE}

# The study site is the world as a whole. Original estimate in USD, but value
# must be converted to euro.

# defining the countries that make up the aggregate
nos <- c("France", "United Kingdom", "Belgium", "Netherlands",
         "Denmark", "Norway", "Germany") # North Sea countries

bt_transfer(study_site = "WLD", policy_site = policy_site, 
            study_yr = 2014, policy_yr = 2020, study_currency = "USD",
            policy_currency = "EUR", aggregate = "yes",
            aggregate_name = "North Sea")


```

In scenario evaluation and cost-benefit analysis might be needed to perform a 
value transfer to a year into the future. The function `bt_transfer` can
accomplish that. The user only has to set `policy_yr` to a year into the future
(up to 2050). GDP per capita and $\epsilon$ for the policy site(s) are estimated
using the the average GDP and GNI per capita growth rates of the last $n$ year,
where $n$ is the number of year between the current year (latest year for which
data are available) and the policy year. For aggregates, GDP and GNI per capita
growth rates are estimated from GNI, GDP and population data and their average
growth rates.

```{r example-future, echo=TRUE}

# The policy year is in the future

bt_transfer(study_site = "WLD", policy_site = policy_site, 
            study_yr = 2014, policy_yr = 2030, study_currency = "USD",
            policy_currency = "EUR", aggregate = "no")


```

## Usage of `compute_sdr`

In cost-benefit analysis and, more in general, in any kind of analysis that
involves the estimation of monetary values referring to a given moment into the
future, it is important to consider the opportunity cost of capital from an
inter-temporal perspective for the society as a whole. In other words, the 
social discount rate (SDR) is a measure of how the society evaluates future
costs/benefits with respect to present ones [@EuropeanCommission2014].
The empirical estimation of the SDR can be performed in several ways, although
the most commonly used is through the formula provided by Ramsey [@ramsey1928].
The function `compute_sdr` accomplishes that and calculates the social discount
rate for any country in the world or aggregate of countries.

```{r example-sdr, echo=TRUE}

# The SDR calculated for a single country

compute_sdr(country = "Italy", policy_yr = 2019, h = 10)

```

The parameter `policy_yr` refers to the year in which the evaluation must take 
place: it cannot be a year into the future. The parameter `h` indicates the 
length of the time window over which computing key parameters.
One of the key parameters in the Ramsey's equation is the the elasticity of
marginal utility of consumption ($\eta$), which can be computed from tax data, 
for the countries for which such data are available. When such data are not
available an average value from the literature is used ($\eta = 1.35$). Such
literature value is used also for the world as a whole. For aggregates $\eta$
is calculated as a weighted average of country-specific $\eta$ using population
data as weights.

```{r example-sdr2, echo=TRUE}

# SDR calculated for the World as a whole

compute_sdr(country = "WLD", policy_yr = 2019, h = 10)


```

```{r example-sdr3, echo=TRUE}

# SDR calculated for an aggregate of countries

compute_sdr(country = nos, policy_yr = 2019, h = 10,
            aggregate = "yes", aggregate_name = "North Sea")

```

```{r example-sdr4, echo=TRUE}

# SDR calculated for the rest of the world, given an aggregate of countries

compute_sdr(country = nos, policy_yr = 2019, h = 10,
            aggregate = "row")

```

# References {-}
